---
title: "Simulation: Craps Simulation"
editor: visual
markdown: 
---

## Simulation (Probability)

The three questions:

1.  Which probability do you want to estimate?

I want to estimate the probability (p) of winning a game after setting a point. Then I want to estimate the probability of winning 7 out of 10 games using this probability.

2.  What is the simulation result (in numbers and visuals)?

    See below

3.  What is the simulation variability?

I want to estimate the variablility in p by repeating the simulation 10 times and determining the mean and variance. 

Make sure to set a seed to make your analysis reproducible.


```{r}
source("craps_simulation.R")

#Question 1 - What is the probability of winning a game after setting a point?

set.seed(20)
# First we simulate multiple games
result <- run_craps_simulation(N = 1000)

# Then we filter games with a point roll
games_with_point <- subset(result, point != "N/A")

# Next we determine the total games with a point set
total_point_games <- nrow(games_with_point)
print(paste("Total games where a point was set:", total_point_games))


# Next we calculate the probability of winning after a point roll to estimate p
wins_after_point <- sum(games_with_point$outcome == "You Win")
wins_after_point
prob_win_after_point <- wins_after_point / total_point_games
print(paste("Probability of winning after a point roll:", round(prob_win_after_point, 4)))
p <- prob_win_after_point

#Finally we estimate the probability using the binomial distribution
binom_prob <- dbinom(7, size = 10, prob = p)
print(paste("Probability of winning exactly 7 out of 10 games:", round(binom_prob, 4)))


```

```{r}

#Question 2 - What is the simulation result and visuals?

# Visualize outcomes for games with point rolls
library(ggplot2)

# Calculate proportions
point_outcomes <- table(games_with_point$outcome)
point_outcomes_df <- as.data.frame(point_outcomes)
point_outcomes_df$Proportion <- point_outcomes_df$Freq / sum(point_outcomes_df$Freq)

# Plot proportions
ggplot(point_outcomes_df, aes(x = Var1, y = Proportion)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Proportions for Games with a Point Set",
    x = "Outcome",
    y = "Proportion"
  ) +
  theme_minimal()

# Visualizing Binomial distribution for n = 10 games
wins <- 0:10
binom_probs <- dbinom(wins, size = 10, prob = p)

binom_data <- data.frame(wins, binom_probs)

ggplot(binom_data, aes(x = wins, y = binom_probs)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Binomial Distribution of Wins in 10 Games",
    x = "Number of Wins",
    y = "Probability"
  ) +
  theme_minimal()


```

```{r}
#Question 3 - What is the simulation variability?

# Simulate 1000 batches of 10 games each
set.seed(20)  # For reproducibility
n_simulations <- 1000
n_games <- 10
simulated_probs <- replicate(n_simulations, {
  games <- run_craps_simulation(N = n_games)
  mean(games$outcome == "You Win")  # Proportion of wins
})

# Plot the variability
simulated_probs_df <- data.frame(simulated_probs)

ggplot(simulated_probs_df, aes(x = simulated_probs)) +
  geom_histogram(binwidth = 0.01, fill = "skyblue", color = "black") +
  labs(
    title = "Simulation Variability of Win Probability",
    x = "Estimated Probability of Winning",
    y = "Frequency"
  ) +
  theme_minimal()



# Mean and standard deviation of the simulated probabilities
mean_prob <- mean(simulated_probs)
sd_prob <- sd(simulated_probs)

print(paste("Mean estimated probability:", round(mean_prob, 4)))
print(paste("Standard deviation of probabilities:", round(sd_prob, 4)))


```
Conclusion: The estimated probability of winning a game after a point set is 0.4117 based on 1000 games. After repeating the simulation 10 times, we get estimate of of the mean probability of 0.496 and a standard deviation of 0.1546. So after a point set, there is a bout a 41% chance of winning a game. If you play 10 games then the probability of winning exactly seven games is 0.049, or only about a 5% chance! 


